---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data
##Introduction

This document refer to the data analysis project required during week 4 of the Linear Regression and Modeling course by Duke University (Coursera.)

The data set is comprised of 651 randomly sampled movies produced and released before 2016.

Some of these variables are only there for informational purposes and do not make any sense to include in a statistical analysis. It is users choice to decide which variables are meaningful and which should be omitted. For example information in the the actor1 through actor5 variables was used to determine whether the movie casts an actor or actress who won a best actor or actress Oscar.




* * *

## Part 2: Research question
Determine which parameters are major predictor for audience_score.
We can generalize the same cocept for other ratings like application user rating,PC game ratings.

####Scope of Inference

For the purposes of inference, this shoud be considered an observational study that uses a random sampling approach to obtain a representative sample from U.S. movies released between 1974 and 2016. Since a random sampling method is applied in data collection, the results can be generalizable to the movies released between 1974 and 2016.

Causation can only be inferred from a randomized experiment. This study does not meet the requirements of a randomized experiment, therefore causation can not be determined.

####Sources of Bias

As Rotten Tomatoes audience score is created by voulnteers, the study may suffer from voluntary response bias since people with strong responses are more likely to participate. The voluntary participants may not be representative of the U.S. population.











* * *

## Part 3: Exploratory data analysis

Below are the variables we have to consider while applying multiple linear regression
audience_score,genre,thtr_rel_month,imdb_rating,critics_score,best_pic_nom,best_actor_win,best_actress_win,top200_box,mpaa_rating,best_dir_win
###Data Pair Plots

Below plot is a plot of paired variables to examine the relationship, if any, that exists between the selected. This method will be used as a quick "eye ball" test ahead of the more in depth model analysis below.


The response variable will be audience_score.

The linear relationship between critics_score and audience_score and imdb_rating and audience_score lead the analyst to believe those data points would have the highest impact on audience score in future lienar models created in this analysis
```{r filter_data}
interestdata<-select(movies,audience_score,genre,thtr_rel_month,imdb_rating,critics_score,best_pic_nom,best_actor_win,best_actress_win,top200_box,mpaa_rating,best_dir_win)
na.exclude(interestdata)
pairs(~audience_score+genre+thtr_rel_month+imdb_rating+critics_score+best_pic_nom+best_actor_win+best_actress_win+top200_box+mpaa_rating+best_dir_win,data=interestdata)
```
* * *

## Part 4: Modeling

For modelling we will use linear regression and start by fitting a model with 11 variables(described in the previous section). Backward elimination will help us to define if better results can be obtained by using a smaller set of attributes. The advantage of backward elimination is that it allows to start with all the variables, deleting one variable at a time until there are no improvements in the model.

here is an initial model with the 11 variables. The adjusted R-squared is 76.68%.
```{r_initial_model}
model<-lm(audience_score~genre+thtr_rel_month+imdb_rating+critics_score+best_pic_nom+best_actor_win+best_actress_win+top200_box+mpaa_rating+best_dir_win,data=interestdata)
summary(model)
```
lets use of the backward functionality to remove model features with low predective value as a means to see if we can come up with a more simple (in terms of number of features) and more effective (same Rsquared or better.)
```{r backword_model}
backwordmodel<-step(model,direction = "backward",trace = FALSE)
summary(backwordmodel)
```
The backwordmodel has only 7 variables and a nearly consistent Adjusted R-squared of 0.7667 , using 2 less variables than the full model. The imdb_rating, genre and the critics score variables are the most significant variables. Mystery and suspens Genre has a strong negative relationship with the reference variable, while Animation genere has a strong positive relationship with the reference variable.

While not sigificantly different than our first model, we will use the backwordmodel.

The Min residuals indicate this model may not be effective when dealing with films with a low audience_score

##Check Conditions

####Multi-collinearity

Per the pairwise plot below, none of the include features appear to share the same or similar relationships with the explanatory variable. Multi-collinearity should not be an issue.

Linear relationship between explanatory and response variables

The strongest linear relationships exist between critics_score and audience_score and imdb_rating and audience_score.
```{r}
m<-lm(formula = audience_score ~ genre + thtr_rel_month + imdb_rating + 
    critics_score + best_pic_nom + best_actress_win, data = interestdata)

pairs(~audience_score + genre + thtr_rel_month + imdb_rating + critics_score + best_pic_nom + best_actress_win,
      data=interestdata)
```
####Nearly Normal Residuals
An observation of the histogram of residuals below indicate a somewhat normal distribution with a strong right skew within the residuals. We see that most of the deviation occurs at the tail of the distriubtion when looking at the Q Q plot. It is clear that this is a linear (not binomial) distribution, so we will consider this test passed.
```{r}
hist(backwordmodel$residuals, main='Histogram of Residuals')
qqnorm(backwordmodel$residuals,main='Normal Probability Plot of Residuals')
qqline(backwordmodel$residuals)
```
####Constant Variability of Residuals
The chart below indicates there is constant variability of residuals
```{r}
plot(backwordmodel$residuals~backwordmodel$fitted,main='Residuals vs. Predicted (fitted) ')
```
####Independent Residuals
The residuals on the chart below seem to be generally homoscedastic. However, there is some degree of heteroscedasticity in the left end of the above visualization; then the model will be less accurate when predicting lower values.
```{r}
plot(backwordmodel$residuals~backwordmodel$fitted,main="Residuals vs. fitted")
abline(0,0)
```

* * *

## Part 5: Prediction

We wanted to predict the audience score for a new movie that has not been used to fit the model. For the movie "Kung Fu Panda 3." The data below, obtained from IMDB and Rotten Tomato represent each respective data point required to populate the SimpleModel.

Impressivley, the model was able to accuratrly created a range of possible values that caputred the actual audience score! The actual audience score for Kung Fu Panda 3 per Rotten Tomatoes was 79, and the model predeicted it to be 83 (rounded down to nearest whole number!) When considering the projected lower and upper boudns of the SimpleModel, we can see the predection falls within the bounds and is very close to the actual score!.
```{r}
KFP3<-data.frame(genre="Animation", thtr_rel_month=1,imdb_rating=7.2,critics_score=87,best_pic_nom="no",best_actress_win="yes")
predict(SimpleModel,KFP3)

```

```{r}
predict(SimpleModel, KFP3, interval="predict") 
```

* * *

## Part 6: Conclusion

The intial research objetvice was Investigate what parameters are major influences on the audience score (audience_score.) Using the SimpleModel, generated with the approach presented above, we were able to identify a 7 feature model that had stastically significant compoenents that accurately were able to predict the audience score of a film from the year 2016 that was not included in the movies data set intially presented.

Potenially using the whole data set, or introducing external factors like net sales adjusted for inflation or social media data, we could have come up with an even more effecitve model, but that can be left the the sequel to this blockbuster!

